---
title: "Data_Viz"
author: "Veronica Cipriani"
date: "22/4/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DATA JOURNALISM PROJECT

The interactive Gmap graph and some plots are computed with Python.  
Some other visualization are, instead, computer in R. The code as follows.

***

##### Linechart for altitude

In order to understand why people decide to go by bus instead of walk to the University, the map with all the different stops is completed by a line chart that shows the precise altitude of each single stops in the route of bus 5.  

```{r}

setwd('')
df <- read.csv('stop_per_linechart.csv', header = TRUE, sep = ',')

library(ggplot2)
library(hrbrthemes)

df$altitude <- as.numeric(df$altitude)
df$stop_name <- as.character(df$stop_name)
df$stop_name <- factor(df$stop_name, levels=unique(df$stop_name))


linechart <- ggplot(df, aes(x=stop_name, y=altitude, group = 1)) + 
  geom_line(aes(color = 'altit. fermate')) +
  geom_area(aes(), alpha = 0.5, position = 'identity') +
  scale_color_manual(values = "#654321") +
  scale_y_continuous(limits = c(150, 400)) +
  labs(title="FERMATE DEL BUS NUMERO 5", 
       subtitle="Line chart che mostra l'altitudine di ogni fermata in ordine di arrivo, 
rispetto all'altezza media (tratteggiata)", 
       caption="Source: GTFS Trento", 
       y="Altitudine (m)", 
       x = "Fermata",
       color=NULL) +
  annotate("text", x = 12.8, y = 185, label = "alt. media", size = 4) +
  geom_hline(yintercept = 194, linetype = 'dashed', color = 'black') +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, size = 8), 
        panel.grid.minor = element_blank()) 


linechart
```
 
 
 ***

##### Linechart for train station departure 

In order to understand the complex situation, a line chart of departure of bus 5 and 5/ is displayed in the article. In addition, the same chart depicts also the arriving number of trains and extra-urban buses, in order to understand also the impact of commuters.  

The plot also underline some hot time-slots, that are due to the hypothetic higher number of passengers in these moments. This is calculus is made on the basis of lectures' timetables.  
In order, we have:

-7 bachelor degree in Povo (each composed of 3 different years with different lectures);  
-10 master degree (each master has 2 years of lectures).  
Given a total number of students of 3665, we can count an approximate number of $=3665/(7*3 + 10* 2) = 90$ students per each lecture. Considering those who live in Povo, take the car or do not attend lectures, a good approximation of students per lectures seems to be 40/45 students.  Once we estimate more or less the number of students per lecture, we need the number of lecture per hour.  

```{r}
setwd('')

library(knitr)
df <- read.csv('table_lezioni.txt', header = TRUE, sep = ',')
kable(df)
```

This count led to consider two slots as the most **crowded**:  
- 8 am to 11 am;  
- 1 pm to 3 pm;  
Therefore, the plots will underline these slots.  

Given this data, the linechart present two yellow areas that underline the most crowded moments during the day.  
```{r}

setwd('')
## ogni 30 min

data <- read.csv('linechart_copia.txt', header = TRUE, sep = ',')
data$X <- NULL

data$h <- as.POSIXct(data$h, format = "%H:%M:%S", tz = "GMT")

library(scales)
library(ggplot2)

ggplot(data, aes(x=h)) + 
  geom_rect(aes(xmin = (h[6]-600), xmax = (h[12]-600), 
                ymin = 0, ymax = 6),
            fill = "gold", alpha = 0.008) +
  geom_rect(aes(xmin = (h[16]-600), xmax = (h[20]-600), 
                ymin = 0, ymax = 6),
            fill = "gold", alpha = 0.008) +
  geom_line(aes(y=count, col=type)) + 
  scale_x_datetime(breaks = date_breaks("2 hour"),
                   labels = date_format("%H:%M")) +
  scale_y_continuous(breaks = round(seq(min(data$count), max(data$count), by = 1),1)) +
  labs(title="LINECHART CON NUMERO DI BUS E TRENI", 
       subtitle="Fermata di Piazza Dante, con evidenziate in giallo
le fasce orarie pi첫 affollate - ogni 30 minuti", 
       caption="Fonti: Trentino Trasporti, Trenitalia", 
       y="Numero di bus/treni (ogni 30 min)", x = 'Ora',
       color=NULL) +
  scale_color_manual(labels = c("bus5", "bus5barra", 'train'), 
                     values = c("bus5" ='#007fff', "bus5barra"= '#00008B', 'train'='#800000')) +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, size = 8),  
        panel.grid.minor = element_blank())  


bus5barra <- sum(data$count[38:44]) # number of buses 5/ in the crowded period in the morning
```



Then, the same plot is represented without the arriving of trains.  

```{r}
setwd('')

data <- read.csv('linechart_copia.txt', header = TRUE, sep = ',')
data <- data[1:44,]
data$X <- NULL

data$h <- as.POSIXct(data$h, format = "%H:%M:%S", tz = "GMT")

library(scales)
library(ggplot2)

ggplot(data, aes(x=h)) + 
  geom_rect(aes(xmin = (h[6]-600), xmax = (h[12]-600), 
                ymin = 0, ymax = 6),
            fill = "gold", alpha = 0.008) +
  geom_rect(aes(xmin = (h[16]-600), xmax = (h[20]-600), 
                ymin = 0, ymax = 6),
            fill = "gold", alpha = 0.008) +
  geom_line(aes(y=count, col=type)) + 
  scale_x_datetime(breaks = date_breaks("2 hour"),
                   labels = date_format("%H:%M")) +
  scale_y_continuous(breaks = round(seq(min(data$count), max(data$count), by = 1),1)) +
  labs(title="LINECHART CON NUMERO DI BUS", 
       subtitle="Fermata di Piazza Dante, con evidenziate in giallo
le fasce orarie pi첫 affollate - ogni 30 minuti", 
       caption="Fonti: Trentino Trasporti", 
       y="Numero di bus (ogni 30 min)", x = 'Ora',
       color=NULL) +
  scale_color_manual(labels = c("bus5", "bus5barra"), 
                     values = c("bus5" ='#007fff', "bus5barra"= '#00008B' )) +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, size = 8),  
        panel.grid.minor = element_blank())  

```

In order to have the same visualization, but with different timespan - 15min - we plot the same data from another dataset.

```{r}
setwd('')
## ogni 15 min 

data <- read.csv('linechart15min.txt', header = TRUE, sep = ',')
data$X <- NULL

data$h <- as.POSIXct(data$h, format = "%H:%M:%S", tz = "GMT")

library(scales)
library(ggplot2)

ggplot(data, aes(x=h)) + 
  geom_rect(aes(xmin = (h[10]+300), xmax = (h[22]+300), 
                ymin = 0, ymax = 4),
            fill = "gold", alpha = 0.005) +
  geom_rect(aes(xmin = (h[30]+300), xmax = (h[38]+300), 
                ymin = 0, ymax = 4),
            fill = "gold", alpha = 0.005) +
  geom_line(aes(y=count, col=type)) + 
  scale_x_datetime(breaks = date_breaks("2 hour"),
                   labels = date_format("%H:%M")) +
  scale_y_continuous(breaks = round(seq(min(data$count), max(data$count), by = 1),1)) +
  labs(title="LINECHART CON NUMERO DI BUS E TRENI", 
       subtitle="Fermata di Piazza Dante, con evidenziate in giallo
le fasce orarie pi첫 affollate - ogni 15 minuti", 
       caption="Fonti: Trentino Trasporti, Trenitalia", 
       y="Numero di bus/treni (ogni 15 min)", x = 'Ora',
       color=NULL) +
  scale_color_manual(labels = c("bus5", "bus5barra", 'train'), 
                     values = c("bus5" ='#007fff', "bus5barra"= '#00008B', 'train'='#800000')) +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, size = 8),  
        panel.grid.minor = element_blank()) 

```

The same plot is also performed with no data about commuters:

```{r}
setwd('')

data <- read.csv('linechart15min.txt', header = TRUE, sep = ',')
data <- data[1:85,]
data$X <- NULL

data$h <- as.POSIXct(data$h, format = "%H:%M:%S", tz = "GMT")

library(scales)
library(ggplot2)

ggplot(data, aes(x=h)) + 
  geom_rect(aes(xmin = (h[10]+300), xmax = (h[22]+300), 
                ymin = 0, ymax = 4),
            fill = "gold", alpha = 0.008) +
  geom_rect(aes(xmin = (h[30]+300), xmax = (h[38]+300), 
                ymin = 0, ymax = 4),
            fill = "gold", alpha = 0.008) +
  geom_line(aes(y=count, col=type)) + 
  scale_x_datetime(breaks = date_breaks("2 hour"),
                   labels = date_format("%H:%M")) +
  scale_y_continuous(breaks = round(seq(min(data$count), max(data$count), by = 1),1)) +
  labs(title="LINECHART CON NUMERO DI BUS", 
       subtitle="Fermata di Piazza Dante, con evidenziate in giallo
le fasce orarie pi첫 affollate - ogni 15 minuti", 
       caption="Fonti: Trentino Trasporti", 
       y="Numero di bus (ogni 15 min)", x = 'Ora',
       color=NULL) +
  scale_color_manual(labels = c("bus5", "bus5barra"), 
                     values = c("bus5" ='#007fff', "bus5barra"= '#00008B' )) +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5, size = 8),  
        panel.grid.minor = element_blank())

```


Given this depicted situation, there is a further 

```{r}
setwd('')

df <- read.csv('table_lezioni.txt', header = TRUE, sep = ',')
dfbus5 <- read.csv('partenze_stazione.txt', header = TRUE, sep = ',')
dfbus5barra <- read.csv('partenze_5barra.txt', header = TRUE, sep = ',')
dfbus5$X <- NULL
dfbus5barra$X <- NULL

dfbus5 <- dfbus5[29:143,]
dfbus5barra <- dfbus5barra[3:34,]

df$ora2 <- as.POSIXct(df$ora, format = "%H:%M:%S", tz = "GMT")
dfbus5$h <- as.POSIXct(dfbus5$departure, format = "%H:%M:%S", tz = "GMT")
dfbus5barra$h <- as.POSIXct(dfbus5barra$departure, format = "%H:%M:%S", tz = "GMT")

d <- (dim(df)[1])
db <- (dim(dfbus5)[1])
dbbarra <- (dim(dfbus5barra)[1])

rownames(dfbus5) <- 1:db
rownames(dfbus5barra) <- 1:dbbarra

df$stima_stud <- df$num.lezioni*40
df$bus5 <- rep(0,dim(df)[1])
df$bus5barra <- rep(0,dim(df)[1])

for (i in 1:d){
  for (j in 1:db){
    if (i == 1) {
      ora <- df$ora2[i]
      oraj <- dfbus5$h[j]
      if (oraj <= ora){
        df$bus5[i] <- df$bus5[i] + dfbus5$partenze.bus[j]
      }
    } else {
      oraj <- dfbus5$h[j]
      prima <- df$ora2[i-1]
      dopo <- df$ora2[i]
      if (prima < oraj && oraj < dopo){
        df$bus5[i] <- df$bus5[i] + dfbus5$partenze.bus[j]
      } else {
        j = j+1
      }
    }
  }
}


for (i in 1:d){
  for (j in 1:dbbarra){
    oraj <- dfbus5barra$h[j]
    if (i == 1) {
      ora <- df$ora2[i]
      if (oraj <= ora){
        df$bus5barra[i] <- df$bus5barra[i] + dfbus5barra$partenze.bus[j]
      }
    } else {
      prima <- df$ora2[i-1]
      dopo <- df$ora2[i]
      if (prima < oraj && oraj < dopo){
        df$bus5barra[i] <- df$bus5barra[i] + dfbus5barra$partenze.bus[j]
      } else {
        j = j+1
      }
    }
  }
}

# per calcolare la portata totale in uno slot orario serve la capienza dei bus
#capienza5 <-
#capienza5barra <-
#df$capienzabus <- df$bus5*capienza5 + df$bus5barra*capienza5barra

df$ora2 <- NULL #per avere solo l'indicazione dell'ora nel df e quindi nella tabella
colnames(df) <- c('Ora', 'Num. Lez.', 'Stima Stud.', 'Bus 5', 'Bus 5/')

write.table(df,"lezioni_studenti.txt",sep=",",row.names=FALSE)

library(knitr)
kable(df)


```
